---
title: "First model"
output: github_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(tidymodels)
library(glmnet)

doParallel::registerDoParallel(core = 4)

load(file = here("data/processed_data.RData"))

playerData <- playerData %>% 
  filter(season < 2022, round <= 23) %>% 
  mutate(brownlowVotes = if_else(brownlowVotes > 0,"Vote","No_vote"),
         full_name = paste(givenName,surname))
```

# First

This is a first model to get a feel for how a predictive model might produce results. Early observations show that a lot of the stats that are often discussed along with Brownlow votes (clearances, goals etc.) tend to be correlated with Brownlow votes. For a first, simple model, Lasso regression will be used to predict the probability of any vote. Votes will be allocated to the players in a given match that have the 3 highest estimated probability of gaining votes in that round, with the highest probability getting 3 votes, second most getting 2 etc. 

## Data splitting

```{r}
brownlowSplit <- initial_split(playerData,prop = 0.75,strata = brownlowVotes)

brownlowTest <- testing(brownlowSplit)
brownlowTrain <- training(brownlowSplit)

brownlowCV <- vfold_cv(brownlowTrain,v = 5)

```

## Set model spec

```{r}
lasso_spec <- logistic_reg(penalty = 0.1, mixture = 1, mode = "classification") %>%
  set_engine("glmnet")
```

## Preprocessing

```{r}
brnlw_rec <- recipe(brownlowVotes ~ centreClearances + goals + totalClearances + stoppageClearances + shotsAtGoal + marksInside50 + goalAssists + ratingPoints + inside50S + goalAccuracy + full_name, data = brownlowTrain) %>% 
  update_role(full_name, new_role = "id")
```

## Workflow

```{r}
wf <- workflow() %>% 
  add_recipe(brnlw_rec) %>% 
  add_model(lasso_spec)
```

## Fit model

```{r}
lasso_fit <- wf %>%
  fit(data = brownlowTrain)

lasso_fit %>%
  pull_workflow_fit() %>%
  tidy()
```

