---
title: "First model"
output: github_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(tidymodels)
library(glmnet)

doParallel::registerDoParallel(core = 4)

load(file = here("data/processed_data.RData"))

playerData_2022 <- playerData %>% 
  filter(season == 2022, round <= 23,!is.na(ratingPoints))

playerData_processed <- playerData %>% 
  filter(season < 2022, round <= 23,!is.na(ratingPoints)) %>% 
  mutate(brownlowVotes_TF = factor(if_else(brownlowVotes > 0,"Vote","No_vote")),
         full_name = paste(givenName,surname)) %>% 
    group_by(season,round,game) %>% 
    mutate(across(c(goals:ruckContests),~dense_rank(desc(.x)),.names = "{.col}_rank")) %>% 
  ungroup() %>% 
  select(-c(startTime:venueName,jumperNumber,captain:timeOnGroundPercentage,contestDefOneOnOnes:kickinsPlayon,umpire1:umpire4,matchup,venueState,Id:ScoreBehinds,extendedStats))
```

# First

This is a first model to get a feel for how a predictive model might produce results. Early observations show that a lot of the stats that are often discussed along with Brownlow votes (clearances, goals etc.) tend to be correlated with Brownlow votes. For a first, simple model, Lasso regression will be used to predict the probability of any vote. Votes will be allocated to the players in a given match that have the 3 highest estimated probability of gaining votes in that round, with the highest probability getting 3 votes, second most getting 2 etc. 

## Data splitting

```{r}
brownlowSplit <- initial_split(playerData_processed,prop = 0.75,strata = brownlowVotes_TF)

brownlowTest <- testing(brownlowSplit)
brownlowTrain <- training(brownlowSplit)

brownlowCV <- vfold_cv(brownlowTrain,v = 5,strata = brownlowVotes_TF)

```


## Preprocessing

```{r}
prep_juice <- function(x) prep(x) %>% juice


brnlw_rec_simple <- recipe(brownlowVotes_TF ~ centreClearances + goals + totalClearances +
                      stoppageClearances + shotsAtGoal + marksInside50 + goalAssists +
                      ratingPoints + inside50S + goalAccuracy + finalMargin + position + full_name,
                    data = brownlowTrain) %>% 
  update_role(full_name,new_role = "id") %>% 
  step_impute_median(all_numeric_predictors()) %>% 
  step_mutate(position = forcats::fct_other(position, keep = c("C","RR","R","FF"))) %>% 
  step_mutate(win = if_else(finalMargin > 0,"Win","Loss") %>% factor) %>% 
  step_rm(finalMargin) %>% 
  step_dummy(position,win) %>% 
  # Upsample brownlow vote rows
  themis::step_smote(brownlowVotes_TF)

brnlw_rec_simple %>% prep_juice

brnlw_rec_extended <- recipe(brownlowVotes_TF ~ goals + behinds + kicks + handballs + disposals + marks + bounces + tackles + contestedPossessions + uncontestedPossessions + totalPossessions + inside50S + marksInside50 + contestedMarks + hitouts + onePercenters + disposalEfficiency + clangers + freesFor + freesAgainst + dreamTeamPoints + rebound50S + goalAssists + goalAccuracy + ratingPoints + turnovers + intercepts + tacklesInside50 + shotsAtGoal + scoreInvolvements + metresGained + centreClearances + stoppageClearances + totalClearances + effectiveKicks + kickEfficiency + effectiveDisposals + marksOnLead + interceptMarks + contestedPossessionRate + hitoutsToAdvantage + hitoutWinPercentage + hitoutToAdvantageRate + groundBallGets + f50GroundBallGets + scoreLaunches + pressureActs + defHalfPressureActs + spoils + ruckContests + finalMargin + position,
                    data = brownlowTrain) %>% 
  step_impute_median(all_numeric_predictors()) %>% 
  step_mutate(position = forcats::fct_other(position, keep = c("C","RR","R","FF"))) %>% 
  step_mutate(win = if_else(finalMargin > 0,"Win","Loss") %>% factor) %>% 
  step_rm(finalMargin) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  # Upsample brownlow vote rows
  themis::step_smote(brownlowVotes_TF)

brnlw_rec_extended %>% prep_juice()

```

## Workflow and model spec

```{r}
lin_wf_rec1 <- workflow() %>%
  add_recipe(brnlw_rec_simple) %>%
  # Add place holders for tunable parameters
  add_model(logistic_reg(penalty = tune(),mixture = tune()) %>%
              set_engine("glmnet"))

lin_wf_rec2 <- workflow() %>%
  add_recipe(brnlw_rec_extended) %>%
  # Add place holders for tunable parameters
  add_model(logistic_reg(penalty = tune(),mixture = tune()) %>%
              set_engine("glmnet"))
```

## Tune hyperparameters

### Recipe 1 - Targeted

```{r}
grid_control <- control_grid(save_pred = TRUE, event_level = "second")

mset <- metric_set(roc_auc,accuracy,specificity,precision)

lin_tune1 <- lin_wf_rec1 %>%
  tune_grid(brownlowCV,
            grid = crossing(penalty = 10 ^ seq(-3, -.5, .1), mixture = c(0.1,0.5,1)),
            metrics = mset,
            control = grid_control)

lin_tune1 %>% 
  autoplot()
```

### Recipe 2 - Kitchen sink

```{r}
lin_tune2 <- lin_wf_rec2 %>%
  tune_grid(brownlowCV,
            grid = crossing(penalty = 10 ^ seq(-3, -.5, .1),mixture = c(0.1,0.5,1)),
            metrics = mset,
            control = grid_control)

lin_tune2 %>% 
  autoplot()
```

### Best fits

```{r}
lin_wf_rec1_tuned <- lin_wf_rec1 %>% 
  finalize_workflow(select_best(lin_tune1,"precision"))

lin_wf_rec2_tuned <- lin_wf_rec2 %>% 
  finalize_workflow(select_best(lin_tune2,"precision"))
```


## Fit resamples

### Recipe 1

```{r}
model_resamples <- 
  as_workflow_set(rec1 = lin_wf_rec1_tuned,rec2 = lin_wf_rec2_tuned) %>% 
    workflow_map("fit_resamples",
                 resamples = brownlowCV,
                 metrics = mset,
                 control = grid_control,
                 verbose = TRUE)

model_resamples %>% 
  collect_metrics()
```

## Best model candidate

```{r}
best_model <- lin_wf_rec2_tuned %>% 
  fit(data = playerData_processed) 

saveRDS(best_model, file = here("models/glm_binary.RDS"))
```

```{r}
glm_coeffs <- best_model %>% 
  extract_fit_parsnip() %>% 
  tidy(exponentiate = TRUE) %>% 
  arrange(-abs(estimate))

glm_coeffs %>% 
  filter(term != "(Intercept)", estimate != 0) %>% 
  ggplot(aes(x = estimate, y = fct_reorder(term,estimate))) + 
  geom_point() +
  geom_vline(xintercept = 0) +
  labs()
```

```{r}
predicted_votes <- playerData_2022 %>% 
  mutate(.pred_Vote = predict(best_model,playerData_2022,type = "prob")$.pred_Vote) %>% 
  group_by(season, round,game) %>% 
  transmute(full_name = paste(givenName,surname),
            teamName,
            brownlowVotes
            ,.pred_Vote,
            game_rank = dense_rank(desc(.pred_Vote)),
            pred_votes = if_else(game_rank > 3L, 0L,game_rank),
            abs_error = abs(brownlowVotes - pred_votes)) %>% 
  ungroup() 
```

```{r}
load(here("data","aflPredictedVotes.RData"))

predicted_votes %>% 
  filter(season == 2022) %>% 
  count(full_name,teamName,wt = pred_votes,sort = T)

predicted_votes %>% 
  filter(season == 2022, full_name == "Touk Miller") %>% 
bind_rows(aflPredictedVotes %>% 
  filter(Player == "Touk Miller") %>% 
  pivot_longer(cols = -Player) %>% 
  transmute(full_name = paste(Player,"AFL"),
         round = str_extract(name,"\\d+"),
         pred_votes = value)) %>% 
  mutate(round = as.integer(round)) %>% 
  ggplot(aes(round,full_name,fill = factor(pred_votes))) +
  geom_tile() +
  geom_text(aes(label = pred_votes)) +
  scale_fill_viridis_d()

```
 

```{r}
predicted_votes %>% 
  filter(season == 2022) %>% 
  transmute(full_name,
            round,
            pred_votes,
            model = "Dave") %>% 
bind_rows(aflPredictedVotes %>% 
            select(-`Round TotalVotes`) %>% 
            pivot_longer(cols = -Player) %>% 
  transmute(full_name = Player,
         round = str_extract(name,"\\d+"),
         pred_votes = value,
         model = "AFL")) %>% 
  count(full_name,model,wt = pred_votes) %>% 
  pivot_wider(names_from = model, values_from = n) %>% 
  arrange(-AFL)
```


## Who was hard done by?

```{r}
playerData %>% 
  mutate(.pred_Vote = predict(model,playerData,type = "prob")$.pred_Vote) %>% 
  filter(brownlowVotes == 0) %>% 
  arrange(-.pred_Vote) %>% 
  select(1:2,game,3:8,
         glm_coeffs %>% 
  filter(term != "(Intercept)") %>% pull(term) %>% any_of)
```

```{r}
playerData %>% 
  count(brownlowVotes, winner = finalMargin > 0) %>% 
  pivot_wider(names_from = winner,values_from = n)
```

```{r}
predict(final_model2,playerData,type = "prob", se.fit = T)
```

