---
title: "Shiny dashboard pre-build exploration"
output: github_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidytext)
library(scales)
library(lubridate)
library(here)

load(file = here("data/processed_data.RData"))

playerData <- playerData %>% 
  filter(season < 2022, round <= 23)
```

# Explore options for a single game dashboard

## Which stats are the most differentiating?

```{r}
playerData %>% 
  mutate(brownlowVotes = if_else(brownlowVotes > 0,"Vote","No_vote")) %>% 
  select(season, brownlowVotes, timeOnGroundPercentage:spoils,
         -c(goalEfficiency, interchangeCounts, shotEfficiency)) %>% 
  pivot_longer(cols = -c(season, brownlowVotes), names_to = "metric") %>% 
  group_by(season,brownlowVotes,metric) %>% 
  summarise(mean = mean(value,na.rm = T),
            .groups = "drop_last") %>% 
  pivot_wider(names_from = brownlowVotes,values_from = mean) %>% 
  mutate(diff = Vote - No_vote,
         pct_diff = log2(Vote/No_vote)) %>% 
  arrange(-abs(pct_diff)) %>% 
  ggplot(aes(x = pct_diff,y = fct_reorder(metric,pct_diff,mean), col = factor(season))) +
  geom_point() + 
  scale_colour_viridis_d() +
  labs(title = "Votes or no votes - who gets more?",
       subtitle = "Difference in mean stat value between vote and no vote cohort \n between 2013-2021",
       caption = "Data: jimmyday12/fizRoy",
       colour = "Season",
       x = "Stat ratio (log2(Vote scored/no votes))",
       y = "Player statistic") +
  theme(plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    plot.caption = element_text(hjust = 1))
```
It seems the clearance stats, and those stats that seem to be associated with gaining possession, or controlled possession are highly valued.

## How many players to visualise? If we look at the metrics with the biggest difference, how many players do we need to compare? Top 5 for each statistic?

 - Need to find a good colour palete
 - Find a good way of distinguishing between teams

```{r}
playerData %>% 
  filter(season == 2021,round == "5",homeTeam == "St Kilda") %>% 
  transmute(season, round, full_name = paste(givenName,surname), teamName, brownlowVotes,
            centreClearances, goals, totalClearances, stoppageClearances, 
         shotsAtGoal, marksInside50, goalAssists, ratingPoints, inside50S, goalAccuracy) %>% 
  pivot_longer(cols = -c(season, round, full_name, teamName,brownlowVotes), names_to = "stat", values_to = "value") %>% 
  group_by(stat) %>% 
  slice_max(value,n = 8, with_ties = FALSE) %>% 
  mutate(full_name = reorder_within(full_name,value,stat)) %>% 
  ggplot(aes(x = value, y = full_name, fill = factor(brownlowVotes))) +
  geom_point(aes(shape = teamName), size = 3,col = "black") +
  scale_y_reordered() +
  #scale_fill_manual(values = c("0" = "grey80", "1" = "#EDF8B1", "2" = "#7FCDBB", "3" = "#2C7FB8")) +
  scale_fill_manual(values = c("grey80","#FC9272", "#FB6A4A", "#A50F15")) +
  scale_shape_manual(values = c(22,24)) +
  labs(title = "Title",
       subtitle = "Sub",
       x = "",
       y = "Player name",
       shape = "Team",
       fill = "Votes") +
  facet_wrap(~stat, scales = "free") +
  theme_bw() +
  guides(fill = guide_legend(override.aes = list(shape = c(21))))
```

## How many ways to summarise the data - raw, % of team/game total, percentile rank?

```{r}
playerData %>% 
  mutate(brownlowVotes = if_else(brownlowVotes > 0,"Vote","No_vote")) %>% 
  select(season, brownlowVotes, 
         centreClearances, goals, totalClearances, stoppageClearances, 
         shotsAtGoal, marksInside50, goalAssists, inside50S, goalAccuracy) %>% 
  pivot_longer(cols = -c(season, brownlowVotes), names_to = "metric") %>% 
  group_by(brownlowVotes, metric) %>% 
  mutate(pct_rank = percent_rank(value)) %>% 
  ggplot(aes(x = pct_rank, y = metric,fill  = brownlowVotes)) +
  ggridges::geom_density_ridges(stat = "binline", alpha = 0.5, scale = 0.9)
```


## Show each vote getters best stats in a given game

```{r}
playerData %>% 
  transmute(season, round, full_name = paste(givenName, surname),
            brownlowVotes, 
         centreClearances, goals, totalClearances, stoppageClearances, 
         shotsAtGoal, marksInside50, goalAssists, inside50S, goalAccuracy) %>% 
  pivot_longer(cols = -c(season, round, full_name, brownlowVotes), names_to = "metric") %>% 
  group_by(season, metric) %>% 
  mutate(pct_rank = percent_rank(value)) %>% 
  filter(brownlowVotes > 0) %>% 
  group_by(full_name, round, season) %>% 
  slice_max(pct_rank,n = 5) %>% 
  filter(str_detect(full_name,"Wines"), season == 2021) %>% 
  ggplot(aes(x = round,y = metric, fill = pct_rank, label = value)) +
  geom_tile() +
  geom_tile(aes(col = factor(brownlowVotes)),size = 1.5, fill = NA) +
  geom_text(col = "white")
```

